Self-sustained carbon monoxide oxidation oscillations on size-selected platinum nanoparticles at atmospheric pressure

### 2012-10-18 paper.tex
Backporting from later versions
Since this document is now checkout from rather long ago
because it fits with pccp, it will be necessary to backport
some newer corrections into this document. Some of this
work has been done in this commit.

\section{Results}
An oscillation measurement is performed by mounting a sample in the setup
directly after cluster deposition and anodic bonding. Initially a light-off CO-
oxidation measurement is performed (shown in
Figure~\ref{fgr:initial_treatment}) where the temperature is increased at
3\,$^\circ$C/min until the sample ignites and achieves full conversion of CO to
CO$_2$. At CO:O$_2$ ratio of $\sim0.08$ and a geometrical Pt coverage of
$\sim0.1\%$ light-off occurs at approximately 180\,$^\circ$C. The temperature
ramp is typically continued up to 260\,$^\circ$C whereafter the temperature is
decreased at the same rate as the heating ramp. After the initial light-off the
sample is cooled to room temperature. When at room temperature the temperature
is again increased at 3\,$^\circ$C/min until sustained oscillations occur. A
typical initial treatment is seen in Figure~\ref{fgr:initial_treatment} where
both the symmetrical light-off ramp and the subsequent increase in temperature
to achieve sustained oscillations is seen. Often the sample will show
oscillations either directly upon light-off or on the falling temperature ramp.

\begin{figure}[h]
\centering
  \includegraphics[width=9cm]{initial_treatment.png}
  \caption{A typical example of the initial treatment of a new sample. After
 performing a light-off ramp where oscillations can been seen the temperature
 is increased to a constant value where sustained oscillations take place. The
 non-zero value of CO during high conversion periods is consistent with the
 expected QMS background signals from CO$_2$ and O$_2$.}
  \label{fgr:initial_treatment}
\end{figure}

Nanoparticles with sizes ranging from 3\,nm to 9\,nm have been tested but 3\,nm
particles with a geometrical coverage of 0.1\% of the reactor area gave the
most stable oscillations. No consistent dependency on oscillation frequency and
duty cycle was found that could be attributed to the size of the particles.

The oscillations on all samples are qualitatively similar but the oscillation
period varies from a few seconds (the time-constant of the reactor is
$\sim7\,$s) up to more than an hour. Once a sample start to oscillate it will
perform self-sustained oscillations for as long as the experiment is allowed to
run. We have not observed a single sample stop oscillating once the self-
sustained oscillations had started.

An example of a single very slow oscillation is seen in
Figure~\ref{fgr:full_oscillation}. After a short on-cycle (at $\sim$1030\,min)
where all CO is converted to CO$_2$ (notice the CO signal does not go to zero
due to the cracking pattern of CO$_2$ and CO background in the QMS) the
reaction shows a fast deactivation. After the deactivation the sample quickly
regains $\sim$30\% conversion (at $\sim$1040\,min). Hereafter the conversion
increases slowly over time until approximately 50\% conversion is reached (at
$\sim$1118\,min). Shortly after reaching 50\% conversion the sample again
ignites and convert all CO to CO$_2$. After a short full conversion period the
sample again deactivates and the cycle repeats itself.

\begin{figure*}
  \centering
  \includegraphics[width=17cm]{single_full_oscillation.png}
  \caption{A full oscillation period first showing a steep ignition of the
  sample followed by an almost immediate deactivation. For the next 65 minutes
  the sample slowly recovers activity until full conversion is reached again
  and the cycle repeats itself. This extremely long oscillation period is not
  commonly seen and was a result of careful parameter tuning. Typical
  oscillation periods are normally between $\sim$30\,s and $\sim$30\,min.
  Inserted circles illustrates the proposed model. At high conversion the
  platinum is oxidized (red) while in low conversion the sample is reduced
  (blue).}
  \label{fgr:full_oscillation}
\end{figure*}

After prolonged measurements over several days an increase in oscillation
period was observed as shown in Figure \ref{fgr:long_measurement}. Initially,
the oscillation period is $\sim$2.5 min and increases linearly with time until
1500 min of total oscillation time. Hereafter the oscillations become more
irregular as shown in Figure~\ref{fgr:extracts} and lower panel
Figure~\ref{fgr:long_measurement}. At experiment end the oscillation period is
between 20 and 45 min.

\begin{figure}[h]
\centering
  \includegraphics[width=9cm]{summary_of_long_measurement.png}
  \caption{Summary of the oscillation period as a function of time for a sample
  oscillating under constant temperature, pressure and reactant composition.
  The sample went through a total of 439 oscillations in 4 days. Time is
  defined from experiment start and hence includes initial treatment. The inset
  shows the initial steady increase in oscillation period.}
  \label{fgr:long_measurement}
\end{figure}

Figure~\ref{fgr:extracts} shows extracts of mass spectrometry data of
oscillations extracted from the same four days measurement. Data from shortly
after the initial treatment, after 2800\,min of oscillation time and after
5900\,min are shown. After 5900\,min the oscillations are more irregular which
is consistent with the data shown in Figure \ref{fgr:long_measurement}. Also,
oscillations in between full conversion oscillations with smaller amplitude and
much higher frequency than the full on-off cycles become more visible as time
progresses.

\begin{figure}[h]
  \centering
  \includegraphics[width=9cm]{extracts_from_very_long_oscillation.png}
  \caption{Extracts from the 4 days long experiment. The oscillation period
  becomes more irregular with time while the total integrated conversion
  remains constant. Furthermore, small oscillations in between full conversion
  cycles become more prominent as time progresses.}
  \label{fgr:extracts}
\end{figure}

@@

%Furthermore, the oscillations have the qualitatively identical behavior as
%shown in Figure~\ref{fgr:full_oscillation}. A short full conversion period is
%followed by a fast deactivation from where reactivity is slowly recovered. At
%approximately 50\% conversion light-off occurs and the sample is in full
%conversion.

\begin{figure}[h]
\centering
  \includegraphics[width=9cm]{temperature_dependence.png}
  \caption{After about 10 hours of oscillations at 205$^\circ$C the temperature
  is first lowered by 5$^\circ$C and soon after increased by 2$^\circ$C. Before
  the temperature step the oscillation period was approximately 500\,s and
  after the step the period is approximately 5000\,s.}
  \label{fgr:temperature_dependence}
\end{figure}

A very strong temperature dependence on the oscillation frequency was also
found as shown in Figure~\ref{fgr:temperature_dependence}. By changing the
temperature 5$^\circ$C the oscillation period was changed by more than a factor
of 10. The magnitude of the change is not completely consistent across all
measured samples but all samples show a very large temperature dependency where
in some cases a provoked small change in temperature will consistently turn on
and off the oscillations.

@@

%No consistent dependency on oscillation frequency and duty cycle was found
%that could be attributed to the size of the particles.

@@

In this study Pt nanoparticles were investigated which, compared to a single
crystal or a thin film, is a very rough surface. It is to be expected that the
particles will initially favor the reduced state. As the reaction runs the gas
composition towards the outlet of the reactor will become more oxidizing
increasing the oxidation rate of the particles (illustrated in
Figure~\ref{fgr:full_oscillation}). Gradually, as the particles toward the
outlet oxidizes the turn-over will increase and the general CO concentration
will decrease promoting more and more particles to oxidize which will gradually
be seen in the QMS as an increase in CO$_2$ and a decrease in CO. As the CO
concentration decreases the light-off temperature will decrease hence
approaching the constant temperature of the sample resulting in a steep
increase in reactivity. It is important to note that the fraction of oxidized
particles needed to achieve full conversion is not known. We are currently
investigating new methods for gaining insight into the nature of the
nanoparticles in the reactor but the low coverage and the design of the reactor
makes this a challenging task. If not all particles are oxidized at light-off
they will now oxidize much faster due to the low CO-concentration. The further
increased rate will not be visible in the QMS since the reactor is already in
full conversion. As most of the particles are now oxidized they will become
more and more roughened and gradually return to the reduced state. Fewer and
fewer oxidized particles will be responsible for the activity increasing the
roughing rate of the remaining active particles. Just before the overall
activity drops only a fraction of the particles participates in the reaction
and the deactivation will happen very suddenly. Once these potentially few
particles eventually becomes sufficiently rough they will reduce and lose
activity hence completing the oscillation cycle. It should be noted that our
group has recently measured turnover frequencies of CO oxidation on Pt in
excess of $10^{6}$s$^{-1}$ per site corresponding to as little as $\sim$300 Pt
nanoparticles contributing to the reactivity just before the end of the high-
activity cycle.

The oscillation period increases with time and becomes increasingly less stable
as shown in Figure~\ref{fgr:extracts}. This phenomena is consistent across all
measured samples. The large scatter in oscillation period can be attributed to
the smaller oscillations in between on/off cycles. Occasionally, these small
amplitude oscillations will trigger a full switch thus introducing more short-
period oscillations despite the trend of increasingly slower oscillation
period. A possible explanation of the phenomeonon could be sintering of the
particles during the oxidation and reductions cycles. However, since the
average conversion integrated over full oscillation periods is constant we do
not expect sintering to be the cause since an average loss in activity of the
sample would then be expected.

@@

\section{Conclusion}
Self-sustained CO oxidation oscillations at atmospheric pressure on Pt
nanoparticles ranging from from 3\,nm to 9\,nm have been measured with high-
quality mass-spectrometry. It was found that the samples increased the
oscillation period over time and showed a strong temperature dependence. A
model is proposed where the nanoparticles go through repetitive
oxidation/reduction cycles which change the change rate of CO oxidation on the
particles.

>>>>

\section{Results}
An oscillation measurement is performed by mounting a sample in the setup directly after cluster deposition and anodic bonding. Initially a light-off CO-oxidation measurement is performed (shown in Figure~\ref{fgr:initial_treatment}) where the temperature is increased at 3\,$^\circ$C/min until the sample ignites and achieves full conversion of CO to CO$_2$. At CO:O$_2$ ratio of $\sim0.08$ and a geometrical Pt coverage of $\sim0.1\%$ light-off occurs at approximately 180\,$^\circ$C. The temperature ramp is typically continued up to 260\,$^\circ$C whereafter the temperature is decreased at the same rate as the heating ramp. After the initial light-off the sample is cooled to room temperature. When at room temperature the temperature is again increased at 3\,$^\circ$C/min until sustained oscillations occur. A typical initial treatment is seen in Figure~\ref{fgr:initial_treatment} where both the symmetrical light-off ramp and the subsequent increase in temperature to achieve sustained oscillations is seen. Often the sample will show oscillations either directly upon light-off or on the falling temperature ramp.
\begin{figure}[h]
\centering
  \includegraphics[width=9cm]{initial_treatment.png}
  \caption{A typical example of the initial treatment of a new sample. After performing a light-off ramp where oscillations can been seen the temperature is increased to a constant value where sustained oscillations take place. The non-zero value of CO during high conversion periods is consistent with the expected QMS background signals from CO$_2$ and O$_2$.}
  \label{fgr:initial_treatment}
\end{figure}
Nanoparticles with sizes ranging from 3\,nm to 9\,nm have been tested but 3\,nm particles with a geometrical coverage of 0.1\% of the reactor area gave the most stable oscillations. No consistent dependency on oscillation frequency and duty cycle was found that could be attributed to the size of the particles.

The oscillations on all samples are qualitatively similar but the oscillation period varies from a few seconds (the time-constant of the reactor is $\sim7\,$s) up to more than an hour. Once a sample start to oscillate it will perform self-sustained oscillations for as long as the experiment is allowed to run. We have not observed a single sample stop oscillating once the self-sustained oscillations had started.

An example of a single very slow oscillation is seen in Figure~\ref{fgr:full_oscillation}. After a short on-cycle (at $\sim$1030\,min) where all CO is converted to CO$_2$ (notice the CO signal does not go to zero due to the cracking pattern of CO$_2$ and CO background in the QMS) the reaction shows a fast deactivation. After the deactivation the sample quickly regains $\sim$30\% conversion (at $\sim$1040\,min). Hereafter the conversion increases slowly over time until approximately 50\% conversion is reached (at $\sim$1118\,min). Shortly after reaching 50\% conversion the sample again ignites and convert all CO to CO$_2$. After a short full conversion period the sample again deactivates and the cycle repeats itself.
\begin{figure*}
  \centering
  \includegraphics[width=17cm]{single_full_oscillation.png}
  \caption{A full oscillation period first showing a steep ignition of the sample followed by an almost immediate deactivation. For the next 65 minutes the sample slowly recovers activity until full conversion is reached again and the cycle repeats itself. This extremely long oscillation period is not commonly seen and was a result of careful parameter tuning. Typical oscillation periods are normally between $\sim$30\,s and $\sim$30\,min. Inserted circles illustrates the proposed model. At high conversion the platinum is oxidized (red) while in low conversion the sample is reduced (blue).}
  \label{fgr:full_oscillation}
\end{figure*}

After prolonged measurements over several days an increase in oscillation period was observed as shown in Figure \ref{fgr:long_measurement}. Initially, the oscillation period is $\sim$2.5 min and increases linearly with time until 1500 min of total oscillation time. Hereafter the oscillations become more irregular as shown in Figure~\ref{fgr:extracts} and lower panel Figure~\ref{fgr:long_measurement}. At experiment end the oscillation period is between 20 and 45 min.
\begin{figure}[h]
\centering
  \includegraphics[width=9cm]{summary_of_long_measurement.png}
  \caption{Summary of the oscillation period as a function of time for a sample oscillating under constant temperature, pressure and reactant composition. The sample went through a total of 439 oscillations in 4 days. Time is defined from experiment start and hence includes initial treatment. The inset shows the initial steady increase in oscillation period.}
  \label{fgr:long_measurement}
\end{figure}

Figure~\ref{fgr:extracts} shows extracts of mass spectrometry data of oscillations extracted from the same four days measurement. Data from shortly after the initial treatment, after 2800\,min of oscillation time and after 5900\,min are shown. After 5900\,min the oscillations are more irregular which is consistent with the data shown in Figure \ref{fgr:long_measurement}. Also, oscillations in between full conversion oscillations with smaller amplitude and much higher frequency than the full on-off cycles become more visible as time progresses.
\begin{figure}[h]
  \centering
  \includegraphics[width=9cm]{extracts_from_very_long_oscillation.png}
  \caption{Extracts from the 4 days long experiment. The oscillation period becomes more irregular with time while the total integrated conversion remains constant. Furthermore, small oscillations in between full conversion cycles become more prominent as time progresses.}
  \label{fgr:extracts}
\end{figure}

@@

%Furthermore, the oscillations have the qualitatively identical behavior as shown in Figure~\ref{fgr:full_oscillation}. A short full conversion period is followed by a fast deactivation from where reactivity is slowly recovered. At approximately 50\% conversion light-off occurs and the sample is in full conversion.

\begin{figure}[h]
\centering
  \includegraphics[width=9cm]{temperature_dependence.png}
  \caption{After about 10 hours of oscillations at 205$^\circ$C the temperature is first lowered by 5$^\circ$C and soon after increased by 2$^\circ$C. Before the temperature step the oscillation period was approximately 500\,s and after the step the period is approximately 5000\,s.}
  \label{fgr:temperature_dependence}
\end{figure}
A very strong temperature dependence on the oscillation frequency was also found as shown in Figure~\ref{fgr:temperature_dependence}. By changing the temperature 5$^\circ$C the oscillation period was changed by more than a factor of 10. The magnitude of the change is not completely consistent across all measured samples but all samples show a very large temperature dependency where in some cases a provoked small change in temperature will consistently turn on and off the oscillations.

@@

%No consistent dependency on oscillation frequency and duty cycle was found that could be attributed to the size of the particles.

@@

In this study Pt nanoparticles were investigated which, compared to a single crystal or a thin film, is a very rough surface. It is to be expected that the particles will initially favor the reduced state. As the reaction runs the gas composition towards the outlet of the reactor will become more oxidizing increasing the oxidation rate of the particles (illustrated in Figure~\ref{fgr:full_oscillation}). Gradually, as the particles toward the outlet oxidizes the turn-over will increase and the general CO concentration will decrease promoting more and more particles to oxidize which will gradually be seen in the QMS as an increase in CO$_2$ and a decrease in CO. As the CO concentration decreases the light-off temperature will decrease hence approaching the constant temperature of the sample resulting in a steep increase in reactivity. It is important to note that the fraction of oxidized particles needed to achieve full conversion is not known. We are currently investigating new methods for gaining insight into the nature of the nanoparticles in the reactor but the low coverage and the design of the reactor makes this a challenging task. If not all particles are oxidized at light-off they will now oxidize much faster due to the low CO-concentration. The further increased rate will not be visible in the QMS since the reactor is already in full conversion. As most of the particles are now oxidized they will become more and more roughened and gradually return to the reduced state. Fewer and fewer oxidized particles will be responsible for the activity increasing the roughing rate of the remaining active particles. Just before the overall activity drops only a fraction of the particles participates in the reaction and the deactivation will happen very suddenly. Once these potentially few particles eventually becomes sufficiently rough they will reduce and lose activity hence completing the oscillation cycle. It should be noted that our group has recently measured turnover frequencies of CO oxidation on Pt in excess of $10^{6}$s$^{-1}$ per site corresponding to as little as $\sim$300 Pt nanoparticles contributing to the reactivity just before the end of the high-activity cycle.

The oscillation period increases with time and becomes increasingly less stable as shown in Figure~\ref{fgr:extracts}. This phenomena is consistent across all measured samples. The large scatter in oscillation period can be attributed to the smaller oscillations in between on/off cycles. Occasionally, these small amplitude oscillations will trigger a full switch thus introducing more short-period oscillations despite the trend of increasingly slower oscillation period. A possible explanation of the phenomeonon could be sintering of the particles during the oxidation and reductions cycles. However, since the average conversion integrated over full oscillation periods is constant we do not expect sintering to be the cause since an average loss in activity of the sample would then be expected.

@@

\section{Conclusion}
Self-sustained CO oxidation oscillations at atmospheric pressure on Pt nanoparticles ranging from from 3\,nm to 9\,nm have been measured with high-quality mass-spectrometry. It was found that the samples increased the oscillation period over time and showed a strong temperature dependence. A model is proposed where the nanoparticles go through repetitive oxidation/reduction cycles which change the change rate of CO oxidation on the particles.
